---
title: "R Notebook"
#output: html_notebook
---

Libraries used
```{r}
library(rio)
library(tidyr)
library(dplyr)
library(stringr)
library(data.table)
library(httr)
library(jsonlite) #install.packages(c("httr", "jsonlite"))
```

```{r}

replaceFile = TRUE # if true, removes files before writing/appending 

delete_file <- function(fn) {
  if (file.exists(fn) && replaceFile) {
    file.remove(fn)
  }
}
```

Import data.
```{r}
import_dat <- import("./output/target_metadata.csv", header=TRUE)
dat <- import_dat
#dat1 <- dat0[,-c(1:2)] # remove extra columns
```

```{r}
dat$count_longname <- str_count(dat$longname, "\\|") + 1 
#categories <- unique(dat$count_longname) #10  1  3  5  4
dat_1 <- dat[dat$count_longname == '1',]
dat_3 <- dat[dat$count_longname == '3',]
dat_4 <- dat[dat$count_longname == '4',]
dat_5 <- dat[dat$count_longname == '5',]
dat_10 <- dat[dat$count_longname == '10',]
```

# Parse Longnames --> Identifiers

Process 10.1: **UniProt > Gene Name**. There are 128 gene names which are not found on PlasmoDB. 
```{r}
new_column_names = c('name', 'transcript', 'gene', 'organism', 'gene_product', 'transcript_product', 'location', 'protein_length', 'sequence_SO', 'SO')
dat_10.2 <- dat_10 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove unnecessary string (e.g. "transcript="). Optional.
unnecessary_string_list = c('transcript', 'gene', 'organism', 'gene_product', 'transcript_product', 'location', 'protein_length', 'sequence_SO', 'SO')
for (val in unnecessary_string_list) {
  dat_10.2[,val] <- gsub(paste0(val,"="), "\\1", dat_10.2[,val])
}

# Remove count_longname column. Optional.
dat_10.3 <- subset(dat_10.2, select=-c(count_longname))

# Make list of identifiers
genes_10 <- list(unique(dat_10.3$gene))

# Write list to file
delete_file("./annotations/genes_10.txt")
invisible(lapply(genes_10, write, "./annotations/genes_10.txt", append=TRUE))
```

Process 5.1: **Sequence Databases > RefSeq Protein**. The second column is version history number? 
```{r}
new_column_names = c('0', '1', 'ref', 'accession', 'definition')
dat_5.2 <- dat_5 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove count_longname column. Optional.
dat_5.3 <- subset(dat_5.2, select=-c(count_longname))

# Make list of gene ids
genes_5 <- list(unique(dat_5.3$accession))

# Write list to file
delete_file("./annotations/genes_5.txt")
invisible(lapply(genes_5, write, "./annotations/genes_5.txt", append=TRUE))
```


Process 4.1: **Not found error... Organism-specific databases > VEuPathDB**. 468 rows are missing identifiers. 
```{r}
new_column_names = c('0', 'identifier', '2','3') 
dat_4.2 <- dat_4 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove count_longname column. Optional.
dat_4.3 <- subset(dat_4.2, select=-c(count_longname))

# Make list of gene ids
genes_4 <- list(unique(dat_4.3$identifier))

# Write list to file
delete_file("./annotations/genes_4.txt")
invisible(lapply(genes_4, write, "./annotations/genes_4.txt", append=TRUE))

# 468 rows are missing identifiers
#dat_4.2 %>%
#  count(identifier)
```

Process 3.1: **UniProt**. Last column can be further parsed? 
```{r}
new_column_names = c('0', 'protein_id', '2') 
dat_3.2 <- dat_3 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove count_longname column. Optional.
dat_3.3 <- subset(dat_3.2, select=-c(count_longname))

# Make list of identifiers
genes_3 <- list(unique(dat_3.3$protein_id))

# Write list to file
delete_file("./annotations/genes_1.txt")
invisible(lapply(genes_3, write, "./annotations/genes_3.txt", append=TRUE))

# Identify Proteins from Uniprot
# https://www.uniprot.org/uniprotkb/

dat_3.2 %>%
  count(protein_id)
```

Process 1.1: 
Measles/Mumps can be found Uniprot or NCBI
PlasmoDB can be found on PlasmoDB or NCBI 
```{r}

# Containing "var" 
dat_1_var <- dat_1[dat_1$longname %like% "var", ]

# Make list of identifiers
genes_1_var <- list(unique(dat_1_var$longname))
# Write list to file
delete_file("./annotations/genes_1_var.txt")
invisible(lapply(genes_1_var, write, "./annotations/genes_1_var.txt", append=TRUE))

##### ####### ###### ###### ###### ###### ######  

# Remaining not containing "var" 
`%notlike%` <- Negate(`%like%`)
dat_1.2 <- dat_1[dat_1$longname %notlike% "var", ]

# Trim for string before the first"_[A-Z]" (underscore followed by a letter/non-number)
dat_1.2$identifier <- sub("\\_[a-zA-Z].*","", dat_1.2$longname)


# Make list of identifiers
genes_1 <- list(unique(dat_1.2$identifier))
# Write list to file
delete_file("./annotations/genes_1.txt")
invisible(lapply(genes_1, write, "./annotations/genes_1.txt", append=TRUE))

```


# Search

UniProt batch retrieve tool: https://www.uniprot.org/id-mapping
Alternatively, programmatically: https://www.uniprot.org/help/programmatic_access
https://www.ncbi.nlm.nih.gov/protein/


Process 10.2: 
```{r}
# info_10 <- import("./annotations/GeneByLocusTag_Summary_10.csv", header=TRUE)
input_ids_10 <- list(unique(info_10$"Input ID"))
not_found_10 <- setdiff(genes_10[[1]],input_ids_10[[1]])
```


Add column for duplicates count. Join by peptide sequence for now. 
```{r}
()
```


