---
title: "R Notebook"
#output: html_notebook
---

Libraries used
```{r}
library(rio)
library(tidyr)
library(dplyr)
library(stringr)
library(data.table)
#library(httr)
#library(jsonlite) #install.packages(c("httr", "jsonlite"))
```

```{r}

replaceFile = TRUE # if true, removes files before writing/appending 

delete_file_if_already_exist <- function(fn) {
  if (file.exists(fn) && replaceFile) {
    file.remove(fn)
  }
}
```

Import data.
```{r}
import_dat <- import("./output/target_metadata.csv", header=TRUE)
dat <- import_dat
#dat1 <- dat0[,-c(1:2)] # remove extra columns
```

```{r}
dat$count_longname <- str_count(dat$longname, "\\|") + 1 
#categories <- unique(dat$count_longname) #10  1  3  5  4
dat_1 <- dat[dat$count_longname == '1',]
dat_3 <- dat[dat$count_longname == '3',]
dat_4 <- dat[dat$count_longname == '4',]
dat_5 <- dat[dat$count_longname == '5',]
dat_10 <- dat[dat$count_longname == '10',]
nrow(dat_1)
nrow(dat_3)
nrow(dat_4)
nrow(dat_5)
nrow(dat_10)
# Remove count_longname column. Optional.
dat_3.3 <- subset(dat_3.2, select=-c(count_longname))
```

# Parse Longnames --> Identifiers

Process 10.1: **Gene Name**
```{r}
new_column_names = c('name', 'transcript', 'gene', 'organism', 'gene_product', 'transcript_product', 'location', 'protein_length', 'sequence_SO', 'SO')
dat_10.2 <- dat_10 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove unnecessary string (e.g. "transcript="). Optional.
unnecessary_string_list = c('transcript', 'gene', 'organism', 'gene_product', 'transcript_product', 'location', 'protein_length', 'sequence_SO', 'SO')
for (val in unnecessary_string_list) {
  dat_10.2[,val] <- gsub(paste0(val,"="), "\\1", dat_10.2[,val])
}

# Write list
gene_names_10_list <- list(unique(dat_10.2$gene))
delete_file_if_already_exist("./annotations/gene_names_10.txt")
invisible(lapply(genenames_10_list, write, "./annotations/gene_names_10.txt", append=TRUE))
```

Process 5.1: **RefSeq Protein**. The second column is version history number? 
```{r}
new_column_names = c('0', '1', 'ref', 'identifier', 'definition')
dat_5.2 <- dat_5 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Write list
refseq_prot_5_list <- list(unique(dat_5.2$identifier))
delete_file_if_already_exist("./annotations/refseq_prot_5.txt")
invisible(lapply(refseq_prot_5_list, write, "./annotations/refseq_prot_5.txt", append=TRUE))
```


Process 4.1: **VEuPathDB:vectorbase:**. Filter out longnames with nonstandard formatting
We filter out the following **manually identified** strings: 
- [blank] 468 rows are missing identifiers. 
- Calvo_et_al
- GI:

We fix: 
- !
- AMEC009308_AMEC013632
```{r}
new_column_names = c('0', 'identifier', '2','3') 
dat_4.2 <- dat_4 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Filter out longnames with nonstandard formatting 
dat_4.2_other <- dat_4.2[dat_4.2$identifier == "" | dat_4.2$longname %like% "Calvo_et_al" | dat_4.2$longname %like% "GI:", ]
# Write list 
delete_file_if_already_exist("./annotations/not_found/list_4_other.txt")
invisible(lapply(list(unique(dat_4.2_other$longname)), write, "./annotations/not_found/list_4_other.txt", append=TRUE))

# Process remaining rows
`%notlike%` <- Negate(`%like%`)
dat_4.3 <- dat_4.2[dat_4.2$identifier != "" & dat_4.2$longname %notlike% "Calvo_et_al" & dat_4.2$longname %notlike% "GI", ]
dat_4.3$identifier <- sub("\\_[a-zA-Z].*","", dat_4.3$identifier) # parse "AMEC009308_AMEC013632" to "AMEC009308"
dat_4.3$identifier <- sub("!","", dat_4.3$identifier) # remove "!"
dat_4.3$vectorbase_identifier <- paste0("vectorbase:", dat_4.3$identifier) # add column for search

# Write list
delete_file_if_already_exist("./annotations/veupathdb_4.txt")
invisible(lapply(list(unique(dat_4.3$vectorbase_identifier)), write, "./annotations/veupathdb_4.txt", append=TRUE))
```

Process 3.1: **UniProt**. Last column can be further parsed? 
```{r}
new_column_names = c('0', 'identifier', '2') 
dat_3.2 <- dat_3 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Write list
uniprot_3_list <- list(unique(dat_3.2$identifier))
delete_file_if_already_exist("./annotations/uniprot_3.txt")
invisible(lapply(uniprot_3_list, write, "./annotations/uniprot_3.txt", append=TRUE))
```

Process 1.1: Filter out longnames with nonstandard formatting
We filter out the following **manually identified** strings: 
- var
- Surfin4.2
- MSPDBL1
- MSPDBL2
- _SG6_reference_seq1
- PFIT_bin
- AMA1
```{r}
# Filter out longnames with nonstandard formatting 
dat_1_other <- dat_1[dat_1$longname %like% "var" | dat_1$longname %like% "Surfin4.2" | dat_1$longname %like% "MSPDBL1" | dat_1$longname %like% "MSPDBL2" | dat_1$longname %like% "_SG6_reference" | dat_1$longname %like% "PFIT_bin"| dat_1$longname %like% "AMA1", ]
# Write list 
delete_file_if_already_exist("./annotations/not_found/list_1_other.txt")
invisible(lapply(list(unique(dat_1_other$longname)), write, "./annotations/not_found/list_1_other.txt", append=TRUE))

# Remaining rows (not filtered out) 
`%notlike%` <- Negate(`%like%`)
dat_1.2 <- dat_1[dat_1$longname %notlike% "var" & dat_1$longname %notlike% "Surfin4.2" & dat_1$longname %notlike% "MSPDBL1" & dat_1$longname %notlike% "MSPDBL2" & dat_1$longname %notlike% "_SG6_reference" & dat_1$longname %notlike% "PFIT_bin" & dat_1$longname %notlike% "AMA1", ]

# Trim for string before the first"_[A-Z]" (underscore followed by a letter/non-number)
dat_1.2$identifier <- sub("\\_[a-zA-Z].*","", dat_1.2$longname)

# Count peptides 
nrow(dat_1_other) # 27156 nonstardard longnames
nrow(dat_1.2) # 6968 remaining peptides
```

Process 1.2: Filter by Species 
Mumps, Human, Enterovirus, Measles, Rubella: 
- Mostly **EMBL-GenBank-DDBJ_CDS** 
P.falciparum strains: 
- Some **EMBL-GenBank-DDBJ** 
- Some **PlasmoDB**
Tetanus, Diptheria, Pertussis: 
- Some **RefSeq** 
```{r}
# P.falciparum 
dat_1_pf <- dat_1.2[dat_1.2$Species == 'P. falciparum',]
# Write list
pf_1_list <- list(unique(dat_1_pf$identifier))
delete_file_if_already_exist("./annotations/pf_1.txt")
invisible(lapply(pf_1_list, write, "./annotations/pf_1.txt", append=TRUE))

# Mumps, Human, Enterovirus, Measles, Rubella
dat_1_embl_prot <- subset(dat_1.2, dat_1.2$Species %in% c('Mumps','Human','Enterovirus','Measles','Rubella'))
# Write list
embl_prot_1_list <- list(unique(dat_1_embl_prot$identifier))
delete_file_if_already_exist("./annotations/embl_prot_1.txt")
invisible(lapply(embl_prot_1_list, write, "./annotations/embl_prot_1.txt", append=TRUE))

# Tetanus, Diptheria, Pertussis
dat_1_refseq_prot <- subset(dat_1.2, dat_1.2$Species %in% c('Tetanus','Diptheria','Pertussis'))
# Write list
refseq_prot_1_list <- list(unique(dat_1_refseq_prot$identifier))
delete_file_if_already_exist("./annotations/refseq_prot_1.txt")
invisible(lapply(refseq_prot_1_list, write, "./annotations/refseq_prot_1.txt", append=TRUE))
```

# Evaluate

Programmatically: 
https://www.uniprot.org/help/id_mapping
https://www.biostars.org/p/9528992/

Load Data
```{r}
gene_names_10 <- import("./annotations/gene_names_10.tsv", header=TRUE)
refseq_prot_5 <- import("./annotations/refseq_prot_5.tsv", header=TRUE)
veupathdb_4 <- import("./annotations/veupathdb_4.tsv", header=TRUE)
uniprot_3 <- import("./annotations/uniprot_3.tsv", header=TRUE)
refseq_prot_1 <- import("./annotations/refseq_prot_1.tsv", header=TRUE)
embl_prot_1 <- import("./annotations/embl_prot_1.tsv", header=TRUE)

```

#Compare

gene_names_10: 4994 hits (4728 unique values). Not found 1591 (previous names/aliases) -> but can be found on **plasmodb.org!**
Finally there is 128 Not found. 
```{r}
gene_names_10_from <- list(unique(gene_names_10$"From"))
gene_names_10_failed <- setdiff(gene_names_10_list[[1]],gene_names_10_from[[1]])

gene_names_10 %>%
  count(From) %>% 
  filter(n > 1)

# Next Step
delete_file_if_already_exist("./annotations/plasmodb_10.txt")
invisible(lapply(gene_names_10_failed, write, "./annotations/plasmodb_10.txt", append=TRUE))

plasmodb_10 <- import("./annotations/plasmodb_10.csv", header=TRUE)
plasmodb_10_from <- list(unique(plasmodb_10$"Input ID"))

# Final Fails. 
all_from_10 <- mapply(c, gene_names_10_from, plasmodb_10_from, SIMPLIFY=FALSE)
all_failed_10 <- setdiff(gene_names_10_list[[1]],all_from_10[[1]])
```

refseq_prot_5: All 17 found (19 hits)
```{r}
refseq_prot_5_from <- list(unique(refseq_prot_5$"From"))
refseq_prot_5_failed <- setdiff(refseq_prot_5_list[[1]],refseq_prot_5_from[[1]])

refseq_prot_5 %>%
  count(From) %>% 
  filter(n > 1)
```

veupathdb_4: 604 hits (439 unique values). Not found 96 (previous names/aliases) -> but can be found on **vectorbase.org!**  
```{r}

veupathdb_4_list <- list(unique(dat_4.3$identifier))
from_output <- list(unique(veupathdb_4$"From")) # correct for string "vectorbase:

veupathdb_4_from <- sub("vectorbase:","", from_output)
veupathdb_4_failed <- setdiff(veupathdb_4_list[[1]],veupathdb_4_from[[1]])

veupathdb_4 %>%
  count(From) %>% 
  filter(n > 1)

# Next Step
delete_file_if_already_exist("./annotations/vectorbase_4.txt")
invisible(lapply(sub("\\vectorbase:","", veupathdb_4_failed), write, "./annotations/vectorbase_4.txt", append=TRUE))

vectorbase_4 <- import("./annotations/vectorbase_4.csv", header=TRUE)
vectorbase_4_from <- list(unique(vectorbase_4$"Input ID"))

# Final Fails. 
all_from_4 <- mapply(c, veupathdb_4_from, vectorbase_4, SIMPLIFY=FALSE)
all_failed_4 <- setdiff(veupathdb_4_list[[1]],all_from_4[[1]])
```

uniprot_3: all 605 found (605 hits)
```{r}
uniprot_3_from <- list(unique(uniprot_3$"From"))
uniprot_3_failed <- setdiff(uniprot_3_list[[1]],uniprot_3_from[[1]])

uniprot_3 %>%
  count(From) %>% 
  filter(n > 1)
```

refseq_prot_1: 10 hits (4 unique values). Not found 4 -> but can be found on **ncbi!**  
```{r}
refseq_prot_1_from <- list(unique(refseq_prot_1$"From"))
refseq_prot_1_failed <- setdiff(refseq_prot_1_list[[1]],refseq_prot_1_from[[1]])

refseq_prot_1 %>%
  count(From) %>% 
  filter(n > 1)
```

embl_prot_1: All 47 found (47 hits)
```{r}
embl_prot_1_from <- list(unique(embl_prot_1$"From"))
embl_prot_1_failed <- setdiff(embl_prot_1_list[[1]],embl_prot_1_from[[1]])

embl_prot_1 %>%
  count(From) %>% 
  filter(n > 1)
```

pf_1: Use **PlasmoDB, Embl_Prot AND Embl**. Not found 5 --> Manually fix --> Not found 2
```{r}
pf_1_plasmodb <- import("./annotations/pf_1_plasmodb.csv", header=TRUE)
pf_1_embl <- import("./annotations/pf_1_embl.tsv", header=TRUE)
pf_1_embl_prot <- import("./annotations/pf_1_embl_prot.tsv", header=TRUE)

pf_1_plasmodb_from <- list(unique(pf_1_plasmodb$"Input ID"))
pf_1_plasmodb_failed <- setdiff(pf_1_list[[1]],pf_1_plasmodb_from[[1]])

pf_1_embl_from <- list(unique(pf_1_embl$"From"))
pf_1_embl_failed <- setdiff(pf_1_list[[1]],pf_1_embl_from[[1]])

pf_1_embl_prot_from <- list(unique(pf_1_embl_prot$"From"))
pf_1_embl_prot_failed <- setdiff(pf_1_list[[1]],pf_1_embl_prot_from[[1]])

# really failed... 
all_from <- mapply(c, pf_1_plasmodb_from, pf_1_embl_from,pf_1_embl_prot_from, SIMPLIFY=FALSE)
all_failed <- setdiff(pf_1_list[[1]],all_from[[1]])
# [1] "PFF0020c"     "PF07_0050_"   "MAL6P1.314"   "PFIT_0801500" "PFIT_0800940"
# Really cannot find are PFIT_0801500 and PFIT_0800940

```


