---
title: "R Notebook"
#output: html_notebook
---

Libraries used
```{r}
library(rio)
library(tidyr)
library(dplyr)
library(stringr)

library(httr)
library(jsonlite) #install.packages(c("httr", "jsonlite"))
```

Import data.
```{r}
import_dat <- import("./output/target_metadata.csv", header=TRUE)
dat <- import_dat
#dat1 <- dat0[,-c(1:2)] # remove extra columns
```

```{r}
dat$count_longname <- str_count(dat$longname, "\\|") + 1 
#categories <- unique(dat$count_longname) #10  1  3  5  4
dat_1 <- dat[dat$count_longname == '1',]
dat_3 <- dat[dat$count_longname == '3',]
dat_4 <- dat[dat$count_longname == '4',]
dat_5 <- dat[dat$count_longname == '5',]
dat_10 <- dat[dat$count_longname == '10',]
```

Process 10.1: After query from PlasmoDB. There are 128 gene ids which are not found on PlasmoDB. 
```{r}
new_column_names = c('name', 'transcript', 'gene', 'organism', 'gene_product', 'transcript_product', 'location', 'protein_length', 'sequence_SO', 'SO')
dat_10.2 <- dat_10 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove unnecessary string (e.g. "transcript="). Optional.
unnecessary_string_list = c('transcript', 'gene', 'organism', 'gene_product', 'transcript_product', 'location', 'protein_length', 'sequence_SO', 'SO')
for (val in unnecessary_string_list) {
  dat_10.2[,val] <- gsub(paste0(val,"="), "\\1", dat_10.2[,val])
}

# Remove count_longname column. Optional.
dat_10.3 <- subset(dat_10.2, select=-c(count_longname))

# Make list of gene ids
genes_10 <- list(unique(dat_10.3$gene))
invisible(lapply(genes_10, write, "./annotations/genes_10.txt", append=TRUE))

# Identify from PlasmoDB. Can also be found NCBI? 
# https://plasmodb.org/plasmo/app/search/transcript/GeneByLocusTag 
```

Process 10.2: 
```{r}
# info_10 <- import("./annotations/GeneByLocusTag_Summary_10.csv", header=TRUE)
input_ids_10 <- list(unique(info_10$"Input ID"))
not_found_10 <- setdiff(genes_10[[1]],input_ids_10[[1]])
```

Process 5.1: 
```{r}
new_column_names = c('0', '1', 'ref', 'accession', 'definition')
dat_5.2 <- dat_5 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove count_longname column. Optional.
dat_5.3 <- subset(dat_5.2, select=-c(count_longname))

# Make list of gene ids
genes_5 <- list(unique(dat_5.3$accession))
invisible(lapply(genes_5, write, "./annotations/genes_5.txt", append=TRUE))

# Identify Proteins from NCBI
# https://www.ncbi.nlm.nih.gov/protein/

```


Process 4.1: 468 rows are missing identifiers
```{r}
new_column_names = c('0', 'identifier', '2','3') 
dat_4.2 <- dat_4 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove count_longname column. Optional.
dat_4.3 <- subset(dat_4.2, select=-c(count_longname))

# Make list of gene ids
genes_4 <- list(unique(dat_4.3$identifier))
invisible(lapply(genes_4, write, "./annotations/genes_4.txt", append=TRUE))

# Identify Proteins from Uniprot: 
# https://www.uniprot.org/uniprotkb/

dat_4.2 %>%
  count(identifier)
```

Process 3.1: Last column can be further parsed? 
```{r}
new_column_names = c('0', 'protein_id', '2') 
dat_3.2 <- dat_3 %>% separate(longname, into = new_column_names, sep = "\\|", remove = FALSE)

# Remove count_longname column. Optional.
dat_3.3 <- subset(dat_3.2, select=-c(count_longname))

# Make list of gene ids
genes_3 <- list(unique(dat_3.3$protein_id))
invisible(lapply(genes_3, write, "./annotations/genes_3.txt", append=TRUE))

# Identify Proteins from Uniprot
# https://www.uniprot.org/uniprotkb/

dat_3.2 %>%
  count(protein_id)
```

Process 1.1: 
Measles/Mumps can be found Uniprot or NCBI
PlasmoDB can be found on PlasmoDB or NCBI 
```{r}
new_column_names = c('0') 
dat_1.2 <- dat_1

# Remove count_longname column. Optional.
dat_1.3 <- subset(dat_1.2, select=-c(count_longname))
```

Add column for duplicates count. Join by peptide sequence for now. 
```{r}

```


